Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-11-25 paper_comparison/paper_comparison/paper_comparison.Rmd
update results

## P-value and power comparison

@@

Figure \ref{fig:power-poly-uniform} shows the estimated power of visual test on lineups with uniform distribution $X_{raw}$ against natural logarithm of the effect with comparison to the power of an exact test - F-test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions. In overall, the power of all four tests increases as the effect becomes larger. The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while others respond inactively to the change of effect and remain lower than 25% throughout the period, showing that as an exact test, the F-test is relatively more sensitive to the type of model defects that being considered. The power of visual test arises steadily and nearly linearly to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features. Other two inappropriate conventional tests shows improvement at the same time but at a lower rate. This coincides the point made by @cook1982residuals mentioned in \ref{hypothesis-testing} that residual-based tests for a specific type of model defect are sensitive to other types of model defects. At $log_e(E) \approx 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin. The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 30% respectively.

What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test. According to the example lineups included in Figure \ref{fig:power-poly-uniform}, at around $log_e(E) = 1$, the  designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model at this level of difficulty. In fact, zero out of 11 subjects detect the actual data plot positioned at panel 10. However, F-test has more than 60% chance of rejecting the null hypothesis in this case. From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public.

The margin is largest at around $log_e(E) = 2$. As shown in Figure \ref{fig:power-poly-uniform}, the visual pattern is still undetectable which is similar to $log_e(E) = 1$. The visual test power is close to zero. In contrast, F-test will reject the null hypothesis 90% of the time.

At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger. At this level of difficulty, the designed visual features are usually detectable but it may not stand out from the lineup as other null plots may happen to include outliers or visual patterns that are are considered to be more attractive by human, and thus recognized as the most different plot. Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison. The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 20, where two of 11 subjects detect it. It can be observed that a "S-shape" is presented in plot 20, but the signal is not strong enough to attract all 11 subjects, resulting in a visual p-value sightly above $0.1$. At $log_e(E) = 4$, the designed visual features become much clear and attractive in panel 3, leading to a high percentage of rejection of the null hypothesis. Figure \ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It is also interesting to observe that the power curve of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible.

@@

for (i in target_sigma) {
  j <- j + 1
  plist[[j]] <- 
  VI_MODEL$plot(poly_model(shape = 2, x = rand_uniform())$
                  gen(n = 100, computed = select(mutate(ex_dat, e = sqrt(i) * e), x, e)), theme = theme_light(base_size =  5), 
              remove_axis = TRUE, 
              remove_grid_line = TRUE) +
    ggtitle(round(es[j], 1))
}

patchwork::wrap_plots(append(plist, list(p), after = 0), 
                      design = map(1:11, ~patchwork::area(11, .x, 11, .x)) %>%
                        reduce(c) %>%
                        c(patchwork::area(1, 1, 10, 11), .))
```

The impact of the distribution of $X_{raw}$ on the power is displayed in Figure \ref{fig:dist-power}. For the non-linearity model, the power curve of F-test is stable across different distributions, while the visual test has a more gentle power curve for the discrete uniform distribution as it contains lineups with larger effect size. BP test performs worse for discrete uniform distribution but has relatively high power for normal distribution. SW test outperforms BP test for discrete uniform distribution but remains as the worst test for other distributions. The results indicate those inappropriate residual-based tests are sensitive to the distribution of the regressor.

For the heteroskedasticity model, F-test is entirely insensitive to the non-constant variance. The visual test has similar power as the SW test  
for normal distribution and lognormal distribution, and has higher power for uniform distribution, and lower power for discrete uniform distribution.

But still, we observe the huge gap in power between the preferred conventional tests and the visual test which is consistent with our previous argument that conventional test is too sensitive for diagnosing non-linearity and heteroskedasticity.

@@

```{r fig.height = 3, fig.width = 12, fig.cap="..."}
patchwork::wrap_plots(append(plist, list(p), after = 0), 
                      design = map(1:11, ~patchwork::area(16, .x, 22, .x)) %>%
                        reduce(c) %>%
                        c(patchwork::area(1, 1, 15, 11), .))
```

@@

```{r fig.height = 3, fig.width = 12, fig.cap="..."}
patchwork::wrap_plots(append(plist, list(p), after = 0), 
                      design = map(1:11, ~patchwork::area(16, .x, 22, .x)) %>%
                        reduce(c) %>%
                        c(patchwork::area(1, 1, 15, 11), .))
```

>>>>

## P-value comparison

@@

## Power comparison

<!-- Figure \ref{fig:power-poly-uniform} shows the estimated power of visual test on lineups with uniform distribution $X_{raw}$ against natural logarithm of the effect with comparison to the power of an exact test - F-test, and two other residual-based conventional tests commonly used in regression diagnostics but for testing other departures from the model assumptions. In overall, the power of all four tests increases as the effect becomes larger. The power curve of F-test climbs aggressively from 25% to around 90% as $log_e(E)$ increases from 0 to 2, while others respond inactively to the change of effect and remain lower than 25% throughout the period, showing that as an exact test, the F-test is relatively more sensitive to the type of model defects that being considered. The power of visual test arises steadily and nearly linearly to around 90% as $log_e(E)$ increases from 2 to 5, suggesting that the effect starts to make noticeable impact on the degree of the presence of the designed visual features. Other two inappropriate conventional tests shows improvement at the same time but at a lower rate. This coincides the point made by @cook1982residuals mentioned in \ref{hypothesis-testing} that residual-based tests for a specific type of model defect are sensitive to other types of model defects. At $log_e(E) \approx 6$, the power curve of F-test reaches almost 100% followed by the visual test by a small margin. The power of Breusch--Pagan test and Shapiro--Wilk test reach around 75% and 30% respectively. -->

<!-- What truly impress us is the huge difference between the estimated power of visual test and the estimated power of F-test. According to the example lineups included in Figure \ref{fig:power-poly-uniform}, at around $log_e(E) = 1$, the  designed visual feature is rarely visible, making the actual data plot indistinguishable from residual plots simulated from the assumed model at this level of difficulty. In fact, zero out of 11 subjects detect the actual data plot positioned at panel 10. However, F-test has more than 60% chance of rejecting the null hypothesis in this case. From a communication perspective, given the fact that the visual difference is unperceivable, the argument that non-linearity present in the fitted model is less convincing to the public. -->

<!-- The margin is largest at around $log_e(E) = 2$. As shown in Figure \ref{fig:power-poly-uniform}, the visual pattern is still undetectable which is similar to $log_e(E) = 1$. The visual test power is close to zero. In contrast, F-test will reject the null hypothesis 90% of the time. -->

<!-- At around $log_e(E) = 3$, the margin gets smaller as the chance of identifying the actual data plot becomes larger. At this level of difficulty, the designed visual features are usually detectable but it may not stand out from the lineup as other null plots may happen to include outliers or visual patterns that are are considered to be more attractive by human, and thus recognized as the most different plot. Without knowing the designed visual features beforehand, it is actually hard to identify the actual data plot by pure image comparison. The corresponding example lineup for $log_e(E) = 3$ shown in Figure \ref{fig:fig:power-poly-uniform} has the actual data plot positioned at panel 20, where two of 11 subjects detect it. It can be observed that a "S-shape" is presented in plot 20, but the signal is not strong enough to attract all 11 subjects, resulting in a visual p-value sightly above $0.1$. At $log_e(E) = 4$, the designed visual features become much clear and attractive in panel 3, leading to a high percentage of rejection of the null hypothesis. Figure \ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It is also interesting to observe that the power curve of visual test behaves similarly to those inappropriate conventional tests until the designed visual features really becomes visible. -->

@@

for (i in target_sigma) {
  j <- j + 1
  plist[[j]] <- 
  VI_MODEL$plot(poly_model(shape = 2, x = rand_uniform())$
                  gen(n = 100, computed = select(mutate(ex_dat, e = sqrt(i) * e), x, e)), theme = theme_light(base_size =  5), 
              remove_axis = TRUE, 
              remove_grid_line = TRUE) +
    ggtitle(round(es[j], 1))
}
```

<!-- The impact of the distribution of $X_{raw}$ on the power is displayed in Figure \ref{fig:dist-power}. For the non-linearity model, the power curve of F-test is stable across different distributions, while the visual test has a more gentle power curve for the discrete uniform distribution as it contains lineups with larger effect size. BP test performs worse for discrete uniform distribution but has relatively high power for normal distribution. SW test outperforms BP test for discrete uniform distribution but remains as the worst test for other distributions. The results indicate those inappropriate residual-based tests are sensitive to the distribution of the regressor.  -->

<!-- For the heteroskedasticity model, F-test is entirely insensitive to the non-constant variance. The visual test has similar power as the SW test   -->
<!-- for normal distribution and lognormal distribution, and has higher power for uniform distribution, and lower power for discrete uniform distribution. -->

<!-- But still, we observe the huge gap in power between the preferred conventional tests and the visual test which is consistent with our previous argument that conventional test is too sensitive for diagnosing non-linearity and heteroskedasticity. -->

@@

```{r polypower, fig.height = 3, fig.width = 12, fig.cap="..."}
patchwork::wrap_plots(append(plist, list(p), after = 0), 
                      design = map(1:11, ~patchwork::area(16, .x, 22, .x)) %>%
                        reduce(c) %>%
                        c(patchwork::area(1, 1, 15, 11), .))
```

@@

```{r heterpower, fig.height = 3, fig.width = 12, fig.cap="..."}
patchwork::wrap_plots(append(plist, list(p), after = 0), 
                      design = map(1:11, ~patchwork::area(16, .x, 22, .x)) %>%
                        reduce(c) %>%
                        c(patchwork::area(1, 1, 15, 11), .))
```

### 2022-11-25 paper_comparison/paper_comparison/paper_comparison.tex
update the result plots
update results

\hypertarget{p-value-and-power-comparison}{%
\subsection{P-value and power
comparison}\label{p-value-and-power-comparison}}

@@

Figure \ref{fig:power-poly-uniform} shows the estimated power of visual
test on lineups with uniform distribution \(X_{raw}\) against natural
logarithm of the effect with comparison to the power of an exact test -
F-test, and two other residual-based conventional tests commonly used in
regression diagnostics but for testing other departures from the model
assumptions. In overall, the power of all four tests increases as the
effect becomes larger. The power curve of F-test climbs aggressively
from 25\% to around 90\% as \(log_e(E)\) increases from 0 to 2, while
others respond inactively to the change of effect and remain lower than
25\% throughout the period, showing that as an exact test, the F-test is
relatively more sensitive to the type of model defects that being
considered. The power of visual test arises steadily and nearly linearly
to around 90\% as \(log_e(E)\) increases from 2 to 5, suggesting that
the effect starts to make noticeable impact on the degree of the
presence of the designed visual features. Other two inappropriate
conventional tests shows improvement at the same time but at a lower
rate. This coincides the point made by \citet{cook1982residuals}
mentioned in \ref{hypothesis-testing} that residual-based tests for a
specific type of model defect are sensitive to other types of model
defects. At \(log_e(E) \approx 6\), the power curve of F-test reaches
almost 100\% followed by the visual test by a small margin. The power of
Breusch--Pagan test and Shapiro--Wilk test reach around 75\% and 30\%
respectively.

What truly impress us is the huge difference between the estimated power
of visual test and the estimated power of F-test. According to the
example lineups included in Figure \ref{fig:power-poly-uniform}, at
around \(log_e(E) = 1\), the designed visual feature is rarely visible,
making the actual data plot indistinguishable from residual plots
simulated from the assumed model at this level of difficulty. In fact,
zero out of 11 subjects detect the actual data plot positioned at panel
10. However, F-test has more than 60\% chance of rejecting the null
hypothesis in this case. From a communication perspective, given the
fact that the visual difference is unperceivable, the argument that
non-linearity present in the fitted model is less convincing to the
public.

The margin is largest at around \(log_e(E) = 2\). As shown in Figure
\ref{fig:power-poly-uniform}, the visual pattern is still undetectable
which is similar to \(log_e(E) = 1\). The visual test power is close to
zero. In contrast, F-test will reject the null hypothesis 90\% of the
time.

At around \(log_e(E) = 3\), the margin gets smaller as the chance of
identifying the actual data plot becomes larger. At this level of
difficulty, the designed visual features are usually detectable but it
may not stand out from the lineup as other null plots may happen to
include outliers or visual patterns that are are considered to be more
attractive by human, and thus recognized as the most different plot.
Without knowing the designed visual features beforehand, it is actually
hard to identify the actual data plot by pure image comparison. The
corresponding example lineup for \(log_e(E) = 3\) shown in Figure
\ref{fig:fig:power-poly-uniform} has the actual data plot positioned at
panel 20, where two of 11 subjects detect it. It can be observed that a
``S-shape'' is presented in plot 20, but the signal is not strong enough
to attract all 11 subjects, resulting in a visual p-value sightly above
\(0.1\). At \(log_e(E) = 4\), the designed visual features become much
clear and attractive in panel 3, leading to a high percentage of
rejection of the null hypothesis. Figure
\ref{fig:fig:power-poly-uniform} gives example lineups of such cases. It
is also interesting to observe that the power curve of visual test
behaves similarly to those inappropriate conventional tests until the
designed visual features really becomes visible.

\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/unnamed-chunk-7-2} \end{center}

The impact of the distribution of \(X_{raw}\) on the power is displayed
in Figure \ref{fig:dist-power}. For the non-linearity model, the power
curve of F-test is stable across different distributions, while the
visual test has a more gentle power curve for the discrete uniform
distribution as it contains lineups with larger effect size. BP test
performs worse for discrete uniform distribution but has relatively high
power for normal distribution. SW test outperforms BP test for discrete
uniform distribution but remains as the worst test for other
distributions. The results indicate those inappropriate residual-based
tests are sensitive to the distribution of the regressor.

For the heteroskedasticity model, F-test is entirely insensitive to the
non-constant variance. The visual test has similar power as the SW
test\\
for normal distribution and lognormal distribution, and has higher power
for uniform distribution, and lower power for discrete uniform
distribution.

But still, we observe the huge gap in power between the preferred
conventional tests and the visual test which is consistent with our
previous argument that conventional test is too sensitive for diagnosing
non-linearity and heteroskedasticity.

>>>>

\hypertarget{p-value-comparison}{%
\subsection{P-value comparison}\label{p-value-comparison}}

@@

\hypertarget{power-comparison}{%
\subsection{Power comparison}\label{power-comparison}}

\begin{landscape}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/polypower-1}

\caption{...}\label{fig:polypower}
\end{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heterpower-1}

\caption{...}\label{fig:heterpower}
\end{figure}

\end{landscape}
